Vagueness, Noise, and Signaling

### 2015-01-15 paper/version_02/paper.tex
rewrite Section 3

The previous section showed that the \rdd can be conceptualized as a
two-step procedure: we first calculate where the replicator dynamic
would take us, then apply diffusion. To understand the \rdd
better, in Section~\ref{sec:iterated-diffusion} we look closer at the novel
diffusion part, before exploring in Section~\ref{sec:simulations} how
diffusion and fitness-based selection interact.

@@

Confusion of states should intuitively be a function of their
perceptual similarity. To make this concrete, let us assume that the
state space of the sim-max game consists of $n \ge 2$ states that are
equally spaced across the unit interval, including $0$ and $1$. The
distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical
similarity between two states $\mystate{i}$ and
$\mystate{j}$. Distance in physical space feeds into a perceptual
similarity function, as described by
\citet{Nosofsky1986:Attention-Simil}:
\begin{align*}
  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \impairment) =
      \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \impairment^2} \right ) & \textrm{otherwise,} \\
    \end{cases}
\end{align*}
where $\impairment >= 0$ is an imprecision or indiscriminability
parameter. When $\impairment=0$ agents perfectly discriminate between
states; when $\impairment \rightarrow \infty$ agents cannot
discriminate states at all. The top part of
Figure~\ref{fig:NosofskySim} gives an impression of
Nosofsky-similarity for different parameter values. Other
formalizations of perceptual similarity are possible, including ones
that allow for different discriminability in different areas of the
state space, but we stick with Nosofsky's similarity function for the
time being, because it is mathematically simple, yet an established
notion in mathematical psychology.

@@

We further assume that the probability of confusing any two states
$\mystate{i}$ and $\mystate{j}$ is proportional to their perceived
similarity and therefore obtained by normalization:
\begin{align*}
  C_{ij} = \frac{\similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} \,.
\end{align*}
Confusion of states is then also a function of imprecision parameter
$\impairment$ (see Figure~\ref{fig:NosofskySim}, bottom part). For $\impairment
= 0$ the confusion matrix has $C_{ii} = 1$; for $\impairment > 0$, $C$ is
positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\impairment
\rightarrow \infty$ we find $C_{ij} = \nicefrac{1}{\card{\States}}$ for all
$i$ and $j$.

If states are confused on average with a probability proportional to
their similarity, repeated applications of diffusion of behavior as
defined in Equation~(\ref{eq:confusion-function}) causes behavioral
dispositions gradually to diffuse along a gradient of similarity of
states as well. Consequently, iterated diffusion leads to a smoothing
out and an eventual equalization of behavioral strategies, for both
the sender and the receiver. After $n$ steps of iterated diffusion an
initial sender strategy $\Sstrat$ will be $C^n \Sstrat$, and an
initial receiver strategy $\Rstrat$ will be $\Rstrat C^n$. If we
assume that $\impairment > 0$, $C$ is a positive row-stochastic matrix
(each state could in principle be confused as any other state, albeit
with possibly a very low probability). It then follows from the
Perron-Frobenius theorem that the limit $C^\infty = \lim_{n
  \rightarrow \infty} C^n$ exists (it is the Perron-projection) and
all of its rows are identical. That is why, in the limit of iterated
diffusion, $C^\infty \Sstrat$ has identical rows (messages are sent with the
same probability in each state), and $\Rstrat C^\infty$ likewise has
identical rows (every message is interpreted equally). All of this is
in line with the intuition that diffusion of behavior under confusion
of states iteratively equalizes behavior for similar states; if all
states can be confused for one another in principle, behavior smoothes
out entirely in the limit.

@@

Now that we understand better what diffusion does, let us look at how
diffusion interacts with optimization of behavior, as described by the
replicator dynamics. The main question we should address is whether
diffusion and fitness-based replication reasonably interact, and if
so, whether the inclusion of diffusion has any noteworthy effects on
the evolving meaning of signals. Clearly, if $\impairment = 0$, the
\rdd reduces to the \rd. If $\impairment \rightarrow \infty$, the
diffusion component takes over and the \rdd reduces to the trivial
iterated diffusion process that we looked at in the previous
section. We will show presently that for reasonable in-between levels
of imprecision $\impairment$, the \rdd leads to communicatively
successful, yet vague signal meaning. Suitable levels of imprecision
can have further accelerating and unifying effects on meaning
evolution.

@@

We ran 50 trials of the \rdd, starting with randomly sampled sender
and receiver strategies, for each triplet of independent parameter
values: $\ns = \card{T} \in \set{6, 10, 50, 90}$, $\impairment \in
\set{0, 0.05, 0.1, 0.2, 0.3}$, $\toler \in \set{0.05, 0.1, 0.2,
  0.3}$. Each trial ran for a maximum of 200 update steps of the
\rdd. A trial was considered converged and stopped before the maximum
of 200 rounds if the total amount of change between strategies before
and after the current \rdd step was smaller than 0.001. It is not
guaranteed that strategies at halting time had converged to the
eventual attracting state, whether they ran for 200 rounds or not. Our
notion of convergence therefore is only a categorical measure for the
speed of approaching some state in which force of evolutionary
optimization has fallen under a (well-considered, but eventually
arbitrary) threshold. In other words, we need to think of convergence
in our sense as speed of evolution towards the first evolutionary
bottleneck.\todo{check and improve this}

Representative examples for resulting strategy pairs at stopping time
are given in
Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA}
shows a strategy pair at stopping time with 90 states, tolerance
$\toler = 0.1$ and imprecision $\impairment = 0$. Zero imprecision
means that the trial was effectively an application of the plain
\rd. Noteworthily, the given sender strategy approximates a pure
sender strategy that crisply partitions the state space into
non-convex sets. The irregular shape of the receiver strategy shows
clearly that the pictured strategy pair has not yet reached a
dynamically stable state under the \rd. Indeed, the trial did not
converge in our technical sense, but was stopped after the maximum of
200 rounds. In contrast, the outcome of a trial with identical
parameters, except with imprecision $\impairment = 0.1$, which is
shown in Figure~\ref{fig:example_stratsB}, had converged (in our
technical sense) after 99 rounds of iterated \rdd. The sender strategy
shows a smooth blending from one ``category'' to the other, and the
receiver's interpretations are rather extended curves, peaking at a
central point in the relevant ``categories.''

@@

To further explore our simulation results, we defined a small number
of metrics that aim to capture numerically how vague, communicatively
efficient and generally well-structured the resulting strategy pairs
were. \emph{Entropy} captures the amount of systematicity or
regularity in signal use. \emph{Convexity} captures whether a
behavioral strategy would project onto a convex pure
strategy. \emph{Expected utility} measures the communicative
efficiency of evolved strategy pairs.\todo{take average of entropy}

@@

The sender strategies in Figures~\ref{fig:example_stratsA} and
\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.21$,
respectively. The receiver strategies have respective entropies $0.43$
and $0.84$. In general, we expect that vague languages will have
higher entropy than crisp ones and that increasing imprecision will
increase entropy, all else equal.

\paragraph{Voronoiness.}
This metric aims to quantify how close a strategy is to being a part
of a Voronoi tessallation of the state space. Bearing in mind that
this metric is specifically geared towards the sim-max games in our
set-up, we define Voronoiness as follows. Let $\pi(\Rstrat, \messg) =
\argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype
of a message $\messg \in \Messgs$, then
\begin{align*}
  V(\Sstrat, \Rstrat) & = \sum_{\state \in \States} \sum_{\messg_1 \in
    \Messgs} \sum_{\messg_2 \in \Messgs} F(\Sstrat, \Rstrat, \state,
  \messg_1, \messg_2) \text{\ \ \ with } \\
  F(\Sstrat, \Rstrat, \state,
  \messg_1, \messg_2) & =
  \begin{cases}
    1 & \text{if } \similarity(\state, \pi(\Rstrat, \messg_1)) >
    \similarity(\state, \pi(\Rstrat, \messg_2)) \bicond \\
    &  \ \ \ \ \ \Sstrat(\state, \messg_1) >
    \Sstrat(\state, \messg_2) \\
    0 & \text{otherwise}
  \end{cases}
\end{align*}
is the Voronoiness of the sender strategy $\Sstrat$ given the receiver
strategy $\Rstrat$ and
\begin{align*}
  V(\Rstrat) & = \sum_{\messg \in
    \Messgs} \sum_{\state_1 \in \States}  \sum_{\state_2 \in \States}
  F(\Rstrat, \messg, \state_1, \state_2) \text{\ \ \ with } \\
  F(\Rstrat, \messg, \state_1, \state_2) & =
  \begin{cases}
    1 & \text{if } \similarity(\state_1, \pi(\Rstrat,
  \messg)) > \similarity(\state_2, \pi(\Rstrat, \messg)) \\
  & \ \ \ \ \ \bicond \Rstrat(\messg,
  \state_1) > \Rstrat(\messg, \state_2) \\
    0 & \text{otherwise}
  \end{cases}
\end{align*}
is the Voronoiness of the receiver strategy $\Rstrat$.  The metrics
are lower bound by $0$ and upper bound by, respectively, $|\States|
\times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$, thus we will
normalize them by dividing by these values.

Despite being vague, a language can have maximal Voronoiness according
to this gradient measure. For instance, the sender strategies in
Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have
Voronoiness $0.91$ and $1$, respectively. The receiver strategies have
Voronoiness $0.78$ and $1$, respectively. We do not \emph{prima facie}
expect a conceptual connection between vagueness and Voronoiness, and
so do not predict a clear effect of imprecision on Voronoiness either.

\paragraph{Convexity.} At least for sender strategies, which converge
faster, it also makes sense to define a categorical measure of
convexity, that compensates for potential vagueness. To determine
whether a sender strategy $\Sstrat$ is convex despite possibly being
vague, we look at the derived pure strategy $\Spure$ for which
$\Spure(\state) = \argmax_{\messg' \in \Messgs}
\Sstrat(\state,\messg')$. If that $\Spure$ is convex, we also count
$\Sstrat$ as convex. The sender strategy in
Figure~\ref{fig:example_stratsA} is not convex, while the one in
Figure~\ref{fig:example_stratsB} is. Again, we do not expect a
conceptual relation between vagueness and imprecision, on the one
hand, and convexity on the other.

\paragraph{Expected utility.} We also record the expected utility of
a strategy pair:
\begin{align*}
  \EU(\Sstrat,\Rstrat \myts ; \myts \toler) = \sum_{\state \in
    \States} \sum_{\messg \in \Messgs} \sum_{\state' \in \States}
  \Pr(\state) \cdot \Sstrat(\state, \messg) \cdot \Rstrat(\messg,
  \state') \cdot \Util(\state,\state' \myts ; \myts \toler)\,.
\end{align*}
To make direct comparisons across different parameter settings, we
normalize expected utility by the maximal amount of expected utility
obtainable in the relevant game.

@@

  \caption{Means of gradient and proportions of categorical metrics for
    each triple of independent variables, with estimated 95\%
    confidence intervals.}
  \label{fig:MeanMetrics}
\end{figure}

Figure~\ref{fig:MeanMetrics} shows the means of the recorded metrics,
together with their estimated 95\% confidence intervals, for each
triple of independent variables. Most trends are quite clear, and not
unexpected. Entropy of sender and receiver strategies are increasing
with imprecision. Entropy of sender strategies seems independent of
tolerance levels and the number of states, while there seems to be an
effect of both parameters on the entropy of the receiver
strategy. What was not expected \emph{a priori} was that Voronoiness
of both sender and receiver strategies appear to be non-decreasing
under increasing impairment, with the only exception being games with
6 states and the transition from $\impairment = 0$ to $\impairment =
0.05$. Expected utility of evolved languages mostly decreases in
increasing imprecision, with the sole exception of a rise for $\toler
= 0.2$ and $n_s = 6$ at $\impairment = 0.1$. Still, normalized
expected utility of evolved strategy pairs under mild values of
imprecision are not necessarily much worse than their respective
perfect precision \rd-analogues. This is especially so for lower
numbers of states (see discussion below).

Taken together, these findings suggest the following effect of
confusability of states and diffusion of behavior. Mild forms of
perceptual imprecision lead to slightly less communicative efficiency,
more vagueness and more regular, well-behaved languages. This
impression is also supported by the proportion of converged and convex
trials, pictured in
Figure~\ref{fig:MeanMetrics}. Non-converged and non-convex
sender strategies showed for several values of parameters in the
absence of impairment. Increasing impairment always assured
convergence and convexity: diffusion speeds up convergence to a convex
and regular language.

% \begin{figure}
%   \centering

%   \includegraphics[width=0.8\textwidth]{plots/CategoricalMeasures.pdf}

%   \caption{Proportion of converged and convex trials.}
%   \label{fig:CategoricalMeasures}
% \end{figure}

Diffusion also has another interesting regularizing effect on the
evolution of signaling. If we look at the confidence intervals in
Figure~\ref{fig:MeanMetrics}, we see that there is very little
variation in the recorded metrics for evolved strategies, at least for
higher values of impairment. On closer inspection, it turns out that
variability in low-impairment conditions is not only due to
non-convergence or non-convexity. Figure~\ref{fig:MoreExample_strats}
gives two more examples of strategy pairs at stopping time. Both are
obtained for the same triple of parameters, both converged and are
convex. However, they are not equally efficient. In fact, the pair in
Figure~\ref{fig:example_stratsC} has a normalized expected utility of
$0.99$ while the one in Figure~\ref{fig:example_stratsD} only has
$0.89$.

@@

Interestingly, this type of variability in evolutionary outcomes is
also weeded out by mild increases in impairment. To see this, we
calculated the average distance between evolved sender strategies
within each group of trials that had identical independent
parameter values. We determined the distance between sender strategies
$\Sstrat$ and $\Sstrat'$ as the average Hellinger distance between
probability distributions $\Sstrat(\state)$ and $\Sstrat'(\state)$
over messages at each choice point $\state$:
\begin{align*}
  \text{HD}(\Sstrat,\Sstrat') & = \frac{1}{{\card{\States} \cdot
     \sqrt{2}}} \cdot  \sum_{\state \in \States} 
 \sqrt{\sum_{\messg \in  \Messgs}
         \left ( \sqrt{\Sstrat(\state,\messg)} -
         \sqrt{\Sstrat'(\state,\messg)} \right )^2 }\,.
\end{align*}
To compensate for the arbitrariness of message use, we set the
distance between strategies $\Sstrat$ and $\Sstrat'$ to be the maximum
of $\text{HD}(\Sstrat,\Sstrat')$ and $\text{HD}(\Sstrat^*,\Sstrat')$
where $\Sstrat^*$ is $\Sstrat$ with reversed message indices. The
\emph{within group distances}, i.e., the average distances between all
sender strategies in each parameter group, are plotted in
Figure~\ref{fig:AverageEUinGroups}. This clearly shows that with
increasing imprecision, the sender strategies evolving under \rdd were
exactly alike (modulo swapping of messages). In other words,
perceptual imprecision unifies evolutionary outcomes, and in fact
guarantees sender strategies that are not only convex, but also
maximally efficient in that they induce a vague category split exactly
in the middle of the unit interval.

@@

This unifying property of perceptual imprecision could be considered
an evolutionary beneficial side-effect. Higher imprecision can lead to
higher \emph{within group expected utility}, i.e., the average
expected utility that each evolved language (i.e., pair of sender and
receiver strategy at stopping time) scored when playing against an
arbitrary other language of the same parameter group (see
Figure~\ref{fig:AverageEUinGroups}). What this means is that, while
imprecision might decrease the communicative efficiency of individual
languages, it increases the conceptual coherence and communicative
success between independently evolving strategies. It is almost as if
mere confusability of states imposes a regularity constraint on
evolving categories.

>>>>

The previous section showed that the \rdd can be conceptualized as a
two-step procedure: we first calculate where the replicator dynamic
would take us, then apply diffusion. To understand the \rdd better, we
look closer at the novel diffusion part in
Section~\ref{sec:iterated-diffusion}, before exploring how diffusion
and fitness-based selection interact in Section~\ref{sec:simulations}.

@@

Confusion of states should intuitively be a function of their
perceptual similarity. To make this concrete, let us assume that the
state space of the sim-max game consists of $n \ge 2$ states that are
equally spaced across the unit interval, including $0$ and $1$. The
distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical
similarity between two states $\mystate{i}$ and
$\mystate{j}$. Distance in physical space feeds into a perceptual
similarity function, as described by
\citet{Nosofsky1986:Attention-Simil}:
\begin{align*}
  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \impairment) =
      \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \impairment^2} \right ) & \textrm{otherwise,} \\
    \end{cases}
\end{align*}
where $\impairment >= 0$ is an imprecision or indiscriminability
parameter. When $\impairment=0$ agents perfectly discriminate between
states; when $\impairment \rightarrow \infty$ agents cannot
discriminate states at all. Figure~\ref{fig:NosofskySim} gives an
impression of Nosofsky-similarity for different parameter
values. Other formalizations of perceptual similarity are possible,
including ones that allow for different discriminability in different
areas of the state space, but we stick with Nosofsky's similarity
function for the time being, because it is mathematically simple, yet
an established notion in mathematical psychology.

@@

We further assume that the probability of confusing any two states
$\mystate{i}$ and $\mystate{j}$ is proportional to their perceived
similarity and therefore obtained by normalization:
\begin{align*}
  C_{ij} = \frac{\similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} \,.
\end{align*}
Confusion of states is then also a function of imprecision parameter
$\impairment$ (see Figure~\ref{fig:NosofskySim}). For $\impairment =
0$ the confusion matrix has $C_{ii} = 1$; for $\impairment > 0$, $C$
is positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\impairment
\rightarrow \infty$ we find $C_{ij} = \nicefrac{1}{\card{\States}}$
for all $i$ and $j$.

If states are confused on average with a probability proportional to
their similarity, repeated application of the operation in
Equation~(\ref{eq:confusion-function}) causes behavioral dispositions
gradually to diffuse along a gradient of similarity of
states. Consequently, iterated diffusion leads to a smoothing out and
an eventual equalization of behavioral strategies, for both the sender
and the receiver. After $n$ steps of iterated diffusion an initial
sender strategy $\Sstrat$ will be $C^n \Sstrat$, and an initial
receiver strategy $\Rstrat$ will be $\Rstrat C^n$. If we assume that
$\impairment > 0$, $C$ is a positive row-stochastic matrix (each state
could in principle be confused as any other state, albeit with
possibly a very low probability). It then follows from the
Perron-Frobenius theorem that the limit $C^\infty = \lim_{n
  \rightarrow \infty} C^n$ exists (it is the Perron-projection) and
all of its rows are identical. That is why, in the limit of iterated
diffusion, $C^\infty \Sstrat$ has identical rows (messages are sent
with the same probability in each state), and $\Rstrat C^\infty$
likewise has identical rows (every message is interpreted
equally). All of this is in line with the intuition that diffusion of
behavior under confusion of states iteratively equalizes behavior for
similar states; if all states can be confused for one another in
principle, behavior smoothes out entirely in the limit.

@@

Now that we understand better what diffusion does, let us look at how
diffusion interacts with optimization of behavior, as described by the
replicator dynamics. The main question we should address is whether
diffusion and fitness-based replication reasonably interact, and if
so, whether the inclusion of diffusion has any noteworthy effects on
the evolving meaning of signals. Clearly, if $\impairment = 0$, the
\rdd reduces to the \rd. If $\impairment \rightarrow \infty$, the
diffusion component takes over and the \rdd reduces to the trivial
iterated diffusion process that we looked at in the previous
section. We will show presently that for reasonable in-between levels
of imprecision $\impairment$, the \rdd can lead to reasonably
efficient, yet vague signaling behavior. Suitable levels of
imprecision can have further accelerating and, surprisingly, unifying
effects on meaning evolution.

@@

We ran 50 trials of the \rdd, starting with randomly sampled sender
and receiver strategies, for each triplet of independent parameter
values: $\ns = \card{T} \in \set{6, 10, 50, 90}$, $\impairment \in
\set{0, 0.05, 0.1, 0.2, 0.3}$, $\toler \in \set{0.05, 0.1, 0.2,
  0.3}$. Each trial ran for a maximum of 200 update steps of the
\rdd. A trial was considered converged and stopped before the maximum
of 200 rounds if the total amount of change between strategies before
and after the current \rdd step was smaller than a suitably chosen
threshold. It is not guaranteed that strategies at halting time had
converged to the eventual attracting state, whether they ran for 200
rounds or not. Our notion of convergence therefore is only a
categorical measure for the speed of approaching some state in which
the force of evolutionary optimization has fallen under a
(well-considered, but eventually arbitrary) threshold. In other words,
we need to think of convergence in our sense as speed of evolution
towards the first evolutionary bottleneck.\todo{check and improve
  this}

Representative examples for resulting strategy pairs are given in
Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA}
shows a strategy pair at stopping time with 90 states, tolerance
$\toler = 0.1$ and imprecision $\impairment = 0$. Zero imprecision
means that the trial was effectively an application of the plain
\rd. Noteworthily, the given sender strategy approximates a pure
sender strategy that crisply partitions the state space into
non-convex sets. The irregular shape of the receiver strategy shows
clearly that the pictured strategy pair has not yet reached a
dynamically stable state under the \rd. Indeed, the trial was stopped
after the maximum of 200 rounds. In contrast, the outcome of a trial
with identical parameters, except with imprecision $\impairment =
0.1$, which is shown in Figure~\ref{fig:example_stratsB}, had
converged (in our technical sense) after 99 rounds of iterated
\rdd. The sender strategy shows a smooth blending from one
``category'' to the other, and the receiver's interpretations are
rather extended curves, peaking at a central point in the relevant
``categories.''

@@

To further explore our simulation results, we track how vague,
communicatively efficient and generally well-structured the recorded
strategy pairs were. \emph{Entropy} captures the amount of
systematicity or regularity in signal use. \emph{Convexity} captures
whether a behavioral strategy would project onto a convex pure
strategy. \emph{Expected utility} measures the communicative
efficiency of evolved strategy pairs.\todo{take average of entropy}

@@

The sender strategies in Figures~\ref{fig:example_stratsA} and
\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.21$,
respectively. The receiver strategies have respective entropies $0.43$
and $0.84$. In general, we expected that vague languages would have
higher entropy than crisp ones and that increasing imprecision would
increase entropy, all else equal.

\paragraph{Convexity.} At least for sender strategies, which converge
faster, it also makes sense to define a categorical measure of
convexity that compensates for potential vagueness. To determine
whether a sender strategy $\Sstrat$ is convex despite possibly being
vague, we look at the derived pure strategy $\Spure$ for which
$\Spure(\state) = \argmax_{\messg' \in \Messgs}
\Sstrat(\state,\messg')$. If that $\Spure$ is convex, we also count
$\Sstrat$ as convex. The sender strategy in
Figure~\ref{fig:example_stratsA} is not convex, while the one in
Figure~\ref{fig:example_stratsB} is.  One would perhaps not expect a
systematic relation between vagueness and imprecision, on the one
hand, and convexity on the other.

\paragraph{Expected utility.} We also recorded the expected utility of
a strategy pair:
\begin{align*}
  \EU(\Sstrat,\Rstrat \myts ; \myts \toler) = \sum_{\state \in
    \States} \sum_{\messg \in \Messgs} \sum_{\state' \in \States}
  \Pr(\state) \cdot \Sstrat(\state, \messg) \cdot \Rstrat(\messg,
  \state') \cdot \Util(\state,\state' \myts ; \myts \toler)\,.
\end{align*}
To make direct comparisons across different parameter settings, we
normalize expected utility by the maximal amount of expected utility
obtainable in the relevant game.

@@

  \caption{Means of gradient and proportions of categorical measures
    for $\toler = 0.1$, $\ns \in \set{6, 10, 50}$, and $\impairment
    \in \set{0, 0.05, 0.1, 0.2, 0.3}$. The plot shows the mean of the
    entropies for the sender and receiver strategy.}
  \label{fig:MeanMetrics}
\end{figure}

Figure~\ref{fig:MeanMetrics} shows plots summarizing the results of
our simulation. For perspicuity, we only plot results for one level of
tolerance $\beta = 0.1$, and leave out the case of $n_s = 90$. Still,
every qualitative trend we mention in the following applies to the
whole set of results.

As expected, increasing imprecision leads to higher entropy and lower
expected utility. Importantly, however, imprecision does not
necessarily lead to disastrous decline of communicative success. On
the other hand, imprecision increased the number of outcomes with
convex sender strategies, and led to higher proportions of
convergence. In fact, sufficient impairment always assured convergence
and convexity. It appears that perceptual imprecision leads to
slightly less communicative efficiency, more vagueness but more
regular, well-behaved languages in shorter time.

Beyond promoting convexity, diffusion also has another interesting
regularizing effect on the evolution of signaling. There was very
little variation in the recorded metrics for evolved strategies, at
least for higher values of impairment. On closer inspection, it turns
out that variability in low-impairment conditions is not only due to
non-convergence or non-convexity. Figure~\ref{fig:MoreExample_strats}
gives two more examples of strategy pairs at stopping time. Both are
obtained for the same triple of parameters, both converged and are
convex. However, they are not equally efficient. In fact, the pair in
Figure~\ref{fig:example_stratsC} has a normalized expected utility of
$0.99$ while the one in Figure~\ref{fig:example_stratsD} only has
$0.89$.

@@

Interestingly, this type of variability in evolutionary outcomes can
be weeded out by impairment. To see this, we calculated the average
distance between evolved sender strategies within each group of trials
that had identical parameter values. We determined the distance
between strategies $\Sstrat$ and $\Sstrat'$ as the average Hellinger
distance between distributions $\Sstrat(\state)$ and
$\Sstrat'(\state)$ at each choice point $\state$:
\begin{align*}
  \text{HD}(\Sstrat,\Sstrat') & = \frac{1}{{\card{\States} \cdot
     \sqrt{2}}} \cdot  \sum_{\state \in \States} 
 \sqrt{\sum_{\messg \in  \Messgs}
         \left ( \sqrt{\Sstrat(\state,\messg)} -
         \sqrt{\Sstrat'(\state,\messg)} \right )^2 }\,.
\end{align*}
To compensate for the arbitrariness of message use, we set the
distance between strategies $\Sstrat$ and $\Sstrat'$ to be the maximum
of $\text{HD}(\Sstrat,\Sstrat')$ and $\text{HD}(\Sstrat^*,\Sstrat')$
where $\Sstrat^*$ is $\Sstrat$ with reversed message indices. The
\emph{within group distances}, i.e., the average distances between all
sender strategies in each parameter group, are plotted exemplarily in
Figure~\ref{fig:AverageEUinGroups} for $\beta = 0.1$ and $n_s =
10$. Despite minor quantitative differences the is the same for all
other parameter settings that we tested: with increasing imprecision,
the sender strategies evolving under \rdd were exactly alike (modulo
swapping of messages). This means that perceptual imprecision unifies
evolutionary outcomes and guarantees sender strategies that are not
only convex, but also regular in that they induce a vague category
split exactly in the middle of the unit interval.

@@

This unifying property of perceptual imprecision could be considered
an evolutionary beneficial side-effect. Higher imprecision can lead to
higher \emph{within group expected utility}, defined as the average
expected utility that each evolved language (i.e., pair of sender and
receiver strategy at stopping time) scored when playing against an
arbitrary other language of the same parameter
group. Figure~\ref{fig:AverageEUinGroups}) gives the representative
example of $\toler = 0.1$ and $n_s = 10$. The observation repeats for
other parameter values: while imprecision might decrease the
communicative efficiency of individual languages, it increases the
conceptual coherence and communicative success between independently
evolving strategies. It is almost as if mere confusability of states
imposes a regularity constraint on evolving categories.
